{\plant}: A package for modelling forest trait ecology \& evolution

### 2015-06-01 ms.tex
Few small edits

% TODO: Might need example/citation for trade-offs here?
The default physiological model used in {\plant} largely reflects that
presented in \citet{Falster-2011,Falster-2015}, but with extensions
allowing for diameter growth also to be estimated. To properly model
trait variation, different parameters (such as rates of leaf turnover,
respiration, photosynthesis per leaf area) in the model are linked via
assumed trade-offs (see Appendix \ref{sec:parameters}).  We refer to
this default model as \texttt{FFW16}, reflecting our initials in this
paper.

@@

While we have implemented one physiological model (\texttt{FFW16}),
the package is designed to allow arbitrary additional physiologies to
be introduced.  That is, rather than varying the parameters of the
equations, the equations themselves can be changed and the higher
level functionality of the package can still be used.  As an
intermediate level, we allow for changes in key parameters to trigger
changes in other parameters (i.e. hyper-parametrisation), allowing
straightforward modelling of trade-offs.  For example, the trait
\textsc{lma} (leaf mass per unit leaf area) is used to estimate the
rate of leaf turnover, based on an observed scaling relationship
spanning across diverse vegetation and plant functional types
\citep{Wright-2004} (see Appendix \ref{sec:FFW16} for details).

@@

Modelling of size-structured dynamics via solving of a deterministic
\textsc{pde} enables multiple demographic phenomena to be
investigated. First, {\plant} can simulate patterns of patch
development after a disturbance, including self-thinning and
successional replacement. Whereas previous approaches for modelling
self-thinning were limited to a single species
\citep[e.g.][]{Barnes-2004, Coomes-2007}, {\plant} accommodates
multiple species with different traits (Fig. \ref{fig:schematic}b;
Fig.  \ref{fig:patch}). Second, {\plant} allows for the effects of
competition and succession on productivity and biomass accumulation
within a patch to be investigated \citep{Falster-2011}. Leaf area
cover and rates of biomass production tend to vary with patch age
\citep{Binkley-2002, Smith-2001, Coomes-2007, Ogawa-2010}. 
% TODO: I really don't know what the last bit of this sentence is
% meant to be saying.  Obviously it's something to do with the
% "emergent" properties of the patch, but not clear to me.
{\plant} enables the effects of competitive feedbacks to be
incorporated, in addition to any intrinsic physiological factors,
yielding emergent non-linear outcomes (Fig. \ref{fig:patch}c).

@@

% TODO: I removed the four(!) references here as I don't think this is
% controversial.
All vegetation is subject to a disturbance regime, such that patches
of the landscape are cleared at some rate, by fire, cyclone od
landslide.  The vegetation thus comprises a series of patches
differing in time since last disturbance and linked via seed dispersal
(Fig.  \ref{fig:schematic}b). Such a system is often referred to as a
structured metapopulation \citep{Gyllenberg-2001}.

@@

Finally, with the extension to the metapopulation level, we are
modelling demography across the entire life cycle and are thus in a
position to estimate invasion fitness -- the ability of a rare
individual with particular traits to establish in a community of
established residents at their equilibrium densities
\citep{Geritz-1998,Dieckmann-2007}.  While ideally calculated as the
long-term per capita growth rate, a more convenient indicator of
fitness in structured metapopulations is given by the
logarithm of the basic reproduction ratio, \(R\)
\citep{Gyllenberg-2001, Metz-2001}.
% TODO: Ideally, don't start a sentence with an abbreviation
\(R\) is simply the total seed output of the new type, averaged across
the metapopulation. So a new type can invade when \(\log R > 0\).

>>>>

The default physiological model used in {\plant} largely reflects that
presented in \citet{Falster-2011,Falster-2015}, but with extensions
allowing for diameter growth also to be estimated. To properly model
trait variation, different parameters (such as rates of leaf turnover,
respiration, photosynthesis per leaf area) in the model are linked via
assumed trade-offs (see Appendix \ref{sec:parameters}).  We refer to
this default model as \texttt{FFW16}, reflecting our initials in this
paper. For example, the trait
\textsc{lma} (leaf mass per unit leaf area) is used to estimate the
rate of leaf turnover, based on an observed scaling relationship
spanning across diverse vegetation and plant functional types
\citep{Wright-2004} (see Appendix \ref{sec:FFW16} for details).

@@

While we have implemented one physiological model (\texttt{FFW16}),
the package is designed to allow arbitrary additional physiologies to
be introduced.  That is, rather than varying the parameters of the
equations, the equations themselves can be changed and the higher
level functionality of the package can still be used.  As an
intermediate level, we allow for changes in key parameters to trigger
changes in other parameters (i.e. hyper-parametrisation), allowing
straightforward modelling of trade-offs.

@@

Modelling of size-structured dynamics via solving of a deterministic
\textsc{pde} enables multiple demographic phenomena to be
investigated. First, {\plant} can simulate demographic 
phenomena typically observed in patches
recovering from a disturbance, including self-thinning and
successional replacement. Whereas previous approaches for modelling
self-thinning were limited to a single species
\citep[e.g.][]{Barnes-2004, Coomes-2007}, {\plant} accommodates
multiple species with different traits (Fig. \ref{fig:schematic}b;
Fig.  \ref{fig:patch}). Moreover, these features emerge from the 
combined effects of physiology and competition for resources, rather 
than being prescribed in the model. Second, {\plant} allows for
the combined effects of physiology and succession on emergent 
characteristics of the entire patch to be investigated
\citep{Falster-2011}. Leaf area cover and rates of biomass 
production tend to vary  non-linearly with patch 
age \citep{Binkley-2002, Smith-2001, Ogawa-2010};
in {\plant} such outcomes arise naturally as emergent properties, 
driven by demography (Fig. \ref{fig:patch}c).

@@

All vegetation is subject to a disturbance regime, such that patches
of the landscape are cleared at some rate, by fire, cyclone od
landslide.  The vegetation thus comprises a series of patches
differing in time since last disturbance and linked via seed dispersal
(Fig.  \ref{fig:schematic}b). Such a system is often referred to as a
structured metapopulation \citep{Gyllenberg-2001}.

@@

Finally, with the extension to the metapopulation level, we are
modelling demography across the entire life cycle and are thus in a
position to estimate invasion fitness -- the ability of a rare
individual with particular traits to establish in a community of
established residents at their equilibrium densities
\citep{Geritz-1998,Dieckmann-2007}.  While ideally calculated as the
long-term per capita growth rate, a more convenient indicator of
fitness in structured metapopulations is given by the
logarithm of the basic reproduction ratio, \(R\)
\citep{Gyllenberg-2001, Metz-2001}.
The basic reproduction ratio is simply the total seed output of 
the new type, averaged across the metapopulation. So a new type can 
invade when \(\log R > 0\).
